---
title: "Trend analysis in time series"
author: "Mark Scheuerell"
output:
  rmarkdown::pdf_document:
    fig_caption: yes        
    includes:  
      in_header: figure_placement.tex
fontsize: 11pt
geometry: margin=1.2in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Trends in time series

Trends in time series can be estimated via simple linear regression where, for a value $x$ measured at time $t$,

$$
x_t = \alpha + \beta t + \epsilon_t.
$$

However, any estimates of the significance of $\alpha$ or $\beta$ will be biased due to non-independence in the observation errors $\epsilon_t$.

## Random walks

In a normal random walk, the value at time $t$ equals that at time $t-1$ plus or minus some random error. Specifically,

\begin{equation} \label{eqnRW}
x_t = x_{t-1} + \epsilon_t.
\end{equation}

Random walks are characterized by long deviations into positive or negative space, but there is no overall tendency to go up or down (Figure 1).

\vspace{0.25in}

```{r rw_plot, echo=FALSE, fig.width=6, fig.height=3, fig.cap="Example of a random walk with Guassian errors."}
set.seed(123)
ww <- rnorm(50)
xx <- cumsum(ww)
plot.ts(xx, ylab=expression(italic(x[t])))
```

# An example



## Requirements

```{r load_up, eval=FALSE}
library(readr)
```


# The data

```{r read_data, eval=FALSE}
## dat <- read_csv()
```


# The analysis

```{r ssm, eval=FALSE}

```

